Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-06-21 PAPER.tex
small edits until 3.2

\item The rate at which empirical eigenvectors converge to their true forms depends on sample variance in the empirical estimate and the steepness of the empirical eigenspectrum (Section \ref{sec:toymodel_frf}). In general, larger sample variances and flatter eigenspectra lead to more loss.

@@

We now apply these lessons to data from the PAPER experiment, using a subset of the PAPER-64 dataset from
\citetalias{ali_et_al2015}) in order to illustrate our revised analysis pipeline.

@@

\noindent of $(22.4$ mK$)^{2}$ in the range $0.15 < k < 0.5$\,$h$ Mpc$^{-1}$ at $z = 8.4$. The need to revise this limit stems from previously under-estimated signal loss and under-estimated error bars, both of which we 
address in the following sections.

For the analysis in this paper, we use $8.1$ hours of LST, namely an RA range of $0.5$-$8.6$ hours (\citetalias{ali_et_al2015} uses a slightly longer RA 
range of $0$-$8.6$ hours; we found that some early LSTs were more severely foreground contaminated). We also use only $10$ baselines, a subset of the $51$ total East/West baselines used in \citetalias{ali_et_al2015}. All power spectrum results are produced for a center frequency of 151\,MHz using a width of 10\,MHz ($20$ channels), identical to the analysis in \citetalias{ali_et_al2015}. In the case study in this paper, we only use one baseline type instead of the three as in 
\citetalias{ali_et_al2015}, but Kolopanis et al. (\textit{in prep.}) uses all baselines presented in \citetalias{ali_et_al2015}.

The most significant changes from \citetalias{ali_et_al2015} occur in our revised power spectrum analysis, which is explained in the rest of this paper, but we also note that the applied fringe-rate filter is also slightly different. In \citetalias{ali_et_al2015}, the 
applied filter was not equivalent to the optimal fringe-rate filter (which is designed to maximize power spectrum sensitivity). Instead, the optimal filter was degraded by widening it in fringe-rate space. This was chosen in order to increase the number of independent 
modes and reduce signal loss, though as we will explain in the next section, this signal loss was still under-estimated. With the development of a new, 
robust method for assessing signal loss, we choose to use the optimal filter in order to maximize sensitivity. This filter is 
computed for a fiducial 30\,m baseline at 150\,MHz, the center frequency in our band. The filter in both the fringe-rate 
domain and time domain is shown in Figure \ref{fig:frp}.

@@

The quantity $\widehat{P}_{\rm in}$, defined by Equation \eqref{eq:Pin}, is a uniformly weighted estimator of the power spectrum of $\mathbf{e}$. Since there is no noise contribution to $\mathbf{e}$, it can be considered the power spectrum of this particular realization of the EoR; alternatively, it can be viewed as the true power spectrum of the injected signal up to cosmic variance fluctuations. The role of $\widehat{P}_{\rm in}$ in our analysis is to serve as a reference for the power spectrum that would be measured if there were no signal loss or other systematics. This is then to be compared to $\widehat{P}_{\rm out}$, which approximates the (lossy) power spectrum estimate that is output by our analysis pipeline prior to any signal loss adjustments. In principle, one could compute the true $\widehat{P}_{\rm out}$ using end-to-end simulations of the instrument and data analysis pipeline. In practice, however, such simulations may not accurately reflect real-life systematics and foregrounds. To overcome this obstacle, one can make the assumption that since the EoR signal is expected to be small, the data vector $\mathbf{x}$ itself is our best model of these contaminants. Making this assumption, the injected EoR signal $\mathbf{e}$ takes on the role of the true EoR signal, and the sum of $\mathbf{x}$ and $\mathbf{e}$ (i.e., Equation \eqref{eq:rxe}) replaces $\mathbf{x}$ as our model of the measured data. Therefore, $\widehat{P}_{\rm out}$ can be directly compared to the measurement that we make.

@@

\begin{eqnarray}
\label{eq:crossterm_full}
\widehat{P}_{\rm out}^{\alpha} &=& \text{M}^{\alpha}_{r}(\textbf{x}+\textbf{e})^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}(\textbf{x}+\textbf{e}) \nonumber \\%- 
%\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x} \nonumber \\
&=& \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} + \text{M}^{\alpha}_{d}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
Assuming \textbf{R}$_{r}$ is symmetric, the two cross-terms (terms with one copy of $\textbf{e}$ and one copy of $\textbf{x}$) can be summed together as:

\begin{eqnarray}
\label{eq:crossterm}
\widehat{P}_{\rm out}^{\alpha} &= &  \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& 2 \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
In order to investigate the effect of each of these terms on signal loss, all three components are plotted in Figure \ref{fig:sigloss_terms} for two cases: empirically estimated inverse covariance weighting ($\textbf{R}_{r} \equiv \widehat{\textbf{C}}_{r}^{-1}$) and uniform weighting ($\textbf{R}_{r} \equiv \textbf{I}$). We will now examine the behavior of this equation in three different regimes of the injected signal - very weak (left ends of the $P_{\rm in}$ axes in Figure \ref{fig:sigloss_terms}), very strong (right ends), and in between (middle portions).

@@

{\bf Large injection:}
When the injected signal is much larger than the measured power spectrum, the data-only components can 
be neglected as many orders of magnitude smaller. We include a description of this regime for completeness in our discussion, but note that the upper limits that we compute are typically not determined by simulations in this regime (i.e. in using an empirical weighting scheme we've assumed the data to be dominated by foregrounds rather than the cosmological signal).  However, it is useful as a check of our system in a relatively simple case. As we can see from Figure \ref{fig:sigloss_terms}, the cross-terms (red) are small in comparison to the signal-only term (green). Here only does the signal-only term used in  \citetalias{ali_et_al2015} dominate the total power output. We again see that, in the estimated inverse covariance weighted case, the cross-terms behave as noise (positive and negative fluctuations around zero mean). This is for the same reason as at small injections --- here $\widehat{\textbf{C}}_{r}$ is dominated by the signal $\textbf{e}$. The cross-correlation can again be modeled as a symmetric distribution of zero mean and width $\sqrt{\widehat{\textbf{P}}_e}\sqrt{\widehat{\textbf{P}}_x}$.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_terms.pdf}
	\caption{Illustration of the power spectrum amplitude of various power spectrum terms as a function of injected EoR power level summed into the 
data. Left: The empirically estimated inverse covariance weighted case used in \citetalias{ali_et_al2015}. The output power spectrum ($\widehat{P}_{\rm out}=\widehat{\textbf{P}}_r$) is shown in black, and the individual terms in Equation 
\eqref{eq:crossterm} are shown in blue, red, and green. The dotted diagonal black line indicates perfect 1:1 input-to-output mapping (no signal loss). The details of the simulation used to generate the figure is explained in Section 
\ref{sec:Practice}; here we sample a larger $P_{\rm in}$ range and fit smooth polynomials to our data points to make an illustrative example.  The gray 
horizontal line is the power spectrum value of data alone, $\widehat{\textbf{P}}_{x}$ (it does not depend on injected power). The green signal-signal 
component is the term used in \citetalias{ali_et_al2015} to estimate signal loss. It is significantly higher 
than $\widehat{\textbf{P}}_{r}$ (black) when the cross-terms (red) are large and negative (black $=$ green $+$ red $+$ blue) . In the 
regime where cross-correlations between injection and data are not dominant (small and large $P_{\rm in}$), the cross-terms have a noise-like 
term with width $\sqrt{\widehat{\textbf{P}}_e}\sqrt{\widehat{\textbf{P}}_x}$. However, at power levels comparable to the data (the middle region), the cross-terms can produce large, negative signal due to the couplings between $\textbf{x}$ and $\textbf{e}$ which affect 
$\widehat{\textbf{C}}_{r}$. This causes the difference between the green curve (which exhibits negligible loss at 
the data-only power spectrum value) and the black curve (which exhibits $\sim$$4$ orders of magnitude of loss). Right: The same power spectrum terms illustrated for the uniform weighted case.}
	\label{fig:sigloss_terms}
\end{figure*}

@@

The loss-corrected power spectrum limit for empirically estimated inverse covariance weighted PAPER-64 data is shown in Figure \ref{fig:ps2_data} (solid red), which we can compare to the original lossy result (dashed red). %There are a few important checks to examine. First, we see that the uniform-weighted power spectrum $2\sigma$ upper limit (dashed blue) is identical in both panels. This is an important check, as we expect no signal loss for this case. Any difference here would indicate an error in the pipeline or discrepancy in the range of data included. Additionally, it is clear that the power spectrum, prior to signal loss estimation, is inconsistent with the (now revised) theoretical noise level prediction. 
Post-signal loss estimation, the power spectrum limits are higher than both the theoretical noise level (green) and uniform-weighted power spectrum (which is shown three ways: black and gray points are positive and negative power spectrum values, respectively, with $2\sigma$ error bars, the solid blue is the upper limit on the EoR signal using the full signal injection framework, and the shaded gray is the power spectrum values with thermal noise errors). We elaborate on this point in the next section, as well as investigate alternate 
weighting schemes to inverse covariance weighting, with the goal of finding one that balances the aggressiveness of down-weighting contaminants and minimizing the loss of the EoR signal.

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data.pdf}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data.pdf}
	\caption{A power spectrum of a subset of PAPER-64 data illustrating the use of empirical inverse covariance weighting. The solid red curve is the $2\sigma$ upper limit on the EoR signal estimated from our signal injection framework using empirical inverse covariance weighting. Shown for comparison is the lossy limit prior to signal loss estimation (dashed red). The theoretical $2\sigma$ thermal noise level prediction based on observational parameters is in green, whose calculation is detailed in Section \ref{sec:PSSense}. Additionally, the power spectrum results for the uniform weighted case is shown in three different ways: power spectrum values (black and gray points as positive and negative values, respectively, with $2\sigma$ error bars), the $2\sigma$ upper limit on the EoR signal using our full signal injection framework (solid blue), and the measured power spectrum values with $2\sigma$ thermal noise errors (gray shaded regions). The vertical dashed black lines signify the horizon limit for this analysis using $30$\,m baselines.}
\label{fig:ps2_data}
\end{figure*}

@@

We saw in Section \ref{sec:otherweight} how limiting the number of down-weighted eigenmodes (i.e. flattening out part of the 
eigenspectrum and effectively de-coupling the lowest-valued eigenmodes from the data, which converge to the true eigenspectrum the slowest and are often EoR-dominated) can help minimize signal loss. We experiment with this idea on PAPER-64 data, dialing the number of modes 
that are down-weighted from zero (which is equivalent to identity-weighting, or the uniform-weighted case) to $21$ (which is the full inverse 
covariance estimator). The power spectrum results for one $k$-value, both before and after signal loss 
estimation, are shown in the top panel in Figure \ref{fig:sigloss_modeloop}. We see that the amount of signal loss increases as weighting 
becomes more aggressive (dashed red). In other words, more EoR-dominated fluctuations are being overfit and 
subtracted as more modes are down-weighted. We also find that the power spectrum upper limit, post-signal loss estimation, 
increases with the number of down-weighted modes (solid red). The more modes we use in down-weighting, the stronger the coupling between the weighting and the data, and the greater the error we have in estimating the power spectrum. \citet{switzer_et_al2013} took a similar approach in determining the optimal number of modes to down-weight in GBT data, finding similar trends and noting that removing too few modes is limited by residual foregrounds and removing too many modes is limited by large error bars and signal loss.

Optimistically, we expect there to be a `sweet spot' as we dial our regularization knob; a level of regularization where weighting 
is beneficial compared to uniform weighting (blue). In other words, we would like a weighting scheme that down-weights eigenmodes that predominantly describe foreground modes, but not EoR modes. We see in Figure \ref{fig:sigloss_modeloop} that this occurs roughly when 
only the $\sim2$-$3$ highest-valued eigenmodes are down-weighted and the rest are given equal weights (though for the case shown, weighting does not actually outperform uniform weighting). For a similar discussion on projecting out modes (zero-ing out eigenmodes, rather than just ignoring their relative weightings as we do in this study), see \citet{switzer_et_al2013}.

We also saw in Section \ref{sec:otherweight} how adding the identity matrix to the empirical covariance can minimize signal loss. We experiment with this idea as well, shown in the bottom panel of Figure \ref{fig:sigloss_modeloop}. The dashed red and solid red lines represent power spectrum limits pre and post-signal loss estimation, respectively, as a function of the strength of $\textbf{I}$ that is added to $\widehat{\textbf{C}}$, quantified as the percentage of Tr($\widehat{\textbf{C}})\textbf{I}$ added to $\widehat{\textbf{C}}$. We parameterize this ``regularization strength" parameter as $\gamma$, namely $\widehat{\textbf{C}} \equiv \widehat{\textbf{C}} + \gamma$Tr$(\widehat{\textbf{C}})\textbf{I}$. From this plot we see that only a small percentage of Tr($\widehat{\textbf{C}})$ is needed to significantly reduce loss. We expect that as the strength of $\textbf{I}$ is increased (going to the left), both the red curves will approach the uniform-weighted case. We also notice that the post-signal loss limit hovers around the uniform-weighted limit for a large range of regularization strengths and while an overall trend from high-to-low signal loss is seen as the strength increases, there does not appear to be a clear `minimum' that produces the least loss.

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_modeloop_2panel.pdf}
	\caption{Power spectra $2\sigma$ upper limits for $k=0.393$\,$h$ Mpc$^{-1}$ for fringe-rate filtered PAPER-64 data. Top: Values 
are shown before (dashed red) and after (solid red) signal loss estimation via our signal injection framework as a function of number of eigenmodes of $\widehat{\textbf{C}}$ that 
are down-weighted. This regularization knob is tuned from $0$ modes on the left (i.e. unweighted) to $21$ modes on the right (i.e. the full inverse 
covariance estimator). Over $\sim3$ orders of magnitude of signal loss results when using empirically estimated inverse covariance weighting. Bottom: Power spectrum upper limits before (dashed red) and after (solid red) signal loss estimation as a function of identity added to the empirical covariance. This regularization knob is tuned from $\gamma = 10^{-4}$ on the right (i.e. very little regularization) to $\gamma = 1$ on the left (see main text for the definition of $\gamma$). Also 
plotted in both panels for comparison are $2\sigma$ power spectrum upper limits for the uniform-weighted case (blue) and inverse variance 
weighted case (black); both are after signal loss estimation. Finally, a theoretical prediction for noise ($2\sigma$ error) is plotted 
as green. In the PAPER-64 analysis in this paper, we choose to use a regularization scheme of $\widehat{\textbf{C}}_{\rm eff} \equiv 0.09 \, $Tr($\widehat{\textbf{C}})\textbf{I} + \widehat{\textbf{C}}$ ($\gamma = 0.09$) as a simple example of regularization that minimizes loss, and note that the power spectrum limits using this type of regularization are roughly constant across a large range of values of $\gamma$.}
	\label{fig:sigloss_modeloop}
\end{figure*}

@@

The power spectrum result for our subset of PAPER-64 data (using only one baseline separation type and $\widehat{\textbf{C}}_{\rm eff}$) is shown in Figure 
\ref{fig:ps1_data}. Again, the solid red curve represents our upper limit on the EoR signal using the full signal injection framework. The uniform weighted case is shown as the black and gray points, which correspond to positive and negative power spectrum values respectively (with 
$2\sigma$ errors bars). It is also shown as an upper limit using the signal injection framework (solid blue), which is interestingly larger than the errors computed from bootstrapping, likely because the full injection framework takes into account additional sample variance whereas the bootstrapped errors do not. Finally, the gray shaded regions combine the measured uniform weighted power spectrum values and the thermal noise. We show this power spectrum result as one example of how a simple regularization can minimize signal loss, though we also note that this weighting does not produce more stringent limits than the uniform weighted case.

>>>>

\item The rate at which empirical eigenvectors converge to their true forms depends on sample variance in the empirical estimate and the steepness (the degeneracy) of the empirical eigenspectrum. In general, larger sample variances and flatter eigenspectra (more degeneracies) lead to more loss (Section \ref{sec:toymodel_frf}).

@@

We now apply these lessons to data from the PAPER experiment, using a subset of the PAPER-64 dataset from
\citetalias{ali_et_al2015} in order to illustrate our revised analysis pipeline.

@@

\noindent of $(22.4$ mK$)^{2}$ in the range $0.15 < k < 0.5$\,$h$ Mpc$^{-1}$ at $z = 8.4$. The need to revise this limit stems mostly from previously under-estimated signal loss and under-estimated error bars, both of which we 
address in the following sections.

For the analysis in this paper, we use $8.1$ hours of LST, namely an RA range of $0.5$-$8.6$ hours (\citetalias{ali_et_al2015} uses a slightly longer RA 
range of $0$-$8.6$ hours; we found that some early LSTs were more severely foreground contaminated). We also use only $10$ baselines, a subset of the $51$ total East/West baselines used in \citetalias{ali_et_al2015}, in order to illustrate our revised methods. All power spectrum results are produced for a center frequency of 151\,MHz using a width of 10\,MHz ($20$ channels), identical to the analysis in \citetalias{ali_et_al2015}. In the case study in this paper, we only use one baseline type instead of the three as in 
\citetalias{ali_et_al2015}, but Kolopanis et al. (\textit{in prep.}) uses the full dataset presented in \citetalias{ali_et_al2015} to revise the result and place limits on the EoR at multiple redshifts.

The most significant changes from \citetalias{ali_et_al2015} occur in our revised power spectrum analysis, which is explained in the rest of this paper, but we also note that the applied fringe-rate filter is also slightly different. In \citetalias{ali_et_al2015}, the 
applied filter was not equivalent to the optimal fringe-rate filter (which is designed to maximize power spectrum sensitivity). Instead, the optimal filter was degraded slightly by widening it in fringe-rate space. This was chosen in order to increase the number of independent 
modes and reduce signal loss, though as we will explain in the next section, this signal loss was still under-estimated. With the development of a new, 
robust method for assessing signal loss, we choose to use the optimal filter in order to maximize sensitivity. This filter is 
computed for a fiducial 30\,m baseline at 150\,MHz, the center frequency in our band. The filter in both the fringe-rate 
domain and time domain is shown in Figure \ref{fig:frp}.

@@

The quantity $\widehat{P}_{\rm in}$, defined by Equation \eqref{eq:Pin}, is a uniformly weighted estimator of the power spectrum of $\mathbf{e}$. Since there is no noise contribution to $\mathbf{e}$, it can be considered the power spectrum of this particular realization of the EoR; alternatively, it can be viewed as the true power spectrum of the injected signal up to cosmic variance fluctuations. The role of $\widehat{P}_{\rm in}$ in our analysis is to serve as a reference for the power spectrum that would be measured if there were no signal loss or other systematics. This is then to be compared to $\widehat{P}_{\rm out}$, which approximates the (lossy) power spectrum estimate that is output by our analysis pipeline prior to any signal loss adjustments. In principle, one could compute the true $\widehat{P}_{\rm out}$ using end-to-end simulations of the instrument and data analysis pipeline. In practice, however, such simulations may not accurately reflect real-life systematics and foregrounds. To overcome this obstacle, one can make the assumption that since the EoR signal is expected to be small, the data vector $\mathbf{x}$ itself is our best model of these contaminants. Making this assumption, the injected EoR signal $\mathbf{e}$ takes on the role of the true EoR signal, and the sum of $\mathbf{x}$ and $\mathbf{e}$ (Equation \eqref{eq:rxe}) replaces $\mathbf{x}$ as our model of the measured data. Therefore, $\widehat{P}_{\rm out}$ can be directly compared to the measurement that we make.

@@

\begin{eqnarray}
\label{eq:crossterm_full}
\widehat{P}_{\rm out}^{\alpha} &=& \text{M}^{\alpha}_{r}(\textbf{x}+\textbf{e})^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}(\textbf{x}+\textbf{e}) \nonumber \\%- 
%\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x} \nonumber \\
&=& \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} + \text{M}^{\alpha}_{d}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x}. %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
Assuming \textbf{R}$_{r}$ is symmetric, the two cross-terms (terms with one copy of $\textbf{e}$ and one copy of $\textbf{x}$) can be summed together as:

\begin{eqnarray}
\label{eq:crossterm}
\widehat{P}_{\rm out}^{\alpha} &= &  \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& 2 \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e}. %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
In order to investigate the effect of each of these terms on signal loss, all three components are plotted in Figure \ref{fig:sigloss_terms} for two cases: empirically estimated inverse covariance weighting ($\textbf{R}_{r} \equiv \widehat{\textbf{C}}_{r}^{-1}$) and uniform weighting ($\textbf{R}_{r} \equiv \textbf{I}$). We will now examine the behavior of this equation in three different regimes of the injected signal - very weak (left ends of the $P_{\rm in}$ axes in Figure \ref{fig:sigloss_terms}), very strong (right ends), and in between (middle portions).

@@

{\bf Large injection:}
When the injected signal is much larger than the measured power spectrum, the data-only components can 
be neglected as many orders of magnitude smaller. We include a description of this regime for completeness in our discussion, but note that the upper limits that we compute are typically not determined by simulations in this regime (i.e. in using an empirical weighting scheme we've assumed the data to be dominated by foregrounds rather than the cosmological signal).  However, it is useful as a check of our system in a relatively simple case. As we can see from Figure \ref{fig:sigloss_terms}, the cross-terms (red) are small in comparison to the signal-only term (green). Here only does the signal-only term used in  \citetalias{ali_et_al2015} dominate the total power output. We again see that, in the empirical inverse covariance weighted case, the cross-terms behave as noise (positive and negative fluctuations around zero mean). This is for the same reason as at small injections --- here $\widehat{\textbf{C}}_{r}$ is dominated by the signal $\textbf{e}$. The cross-correlation can again be modeled as a symmetric distribution of zero mean and width $\sqrt{\widehat{\textbf{P}}_e}\sqrt{\widehat{\textbf{P}}_x}$.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_terms.pdf}
	\caption{Illustration of the power spectrum amplitude of different power spectrum terms as a function of injected EoR power level summed into the 
data. Left: The empirically estimated inverse covariance weighted case used in \citetalias{ali_et_al2015}. The output power spectrum ($\widehat{P}_{\rm out}=\widehat{\textbf{P}}_r$) is shown in black, and the individual terms in Equation 
\eqref{eq:crossterm} are shown in blue, red, and green. The dotted diagonal black line indicates perfect 1:1 input-to-output mapping (no signal loss). The details of the simulation used to generate the figure is explained in Section 
\ref{sec:Practice}; here we sample a larger $P_{\rm in}$ range and fit smooth polynomials to our data points to make an illustrative example.  The gray 
horizontal line is the power spectrum value of data alone, $\widehat{\textbf{P}}_{x}$ (it does not depend on injected power). The green signal-signal 
component is the term used in \citetalias{ali_et_al2015} to estimate signal loss. It is significantly higher 
than $\widehat{\textbf{P}}_{r}$ (black) when the cross-terms (red) are large and negative (black $=$ green $+$ red $+$ blue) . In the 
regime where cross-correlations between injection and data are not dominant (small and large $P_{\rm in}$), the cross-terms have a noise-like 
term with width $\sqrt{\widehat{\textbf{P}}_e}\sqrt{\widehat{\textbf{P}}_x}$. However, at power levels comparable to the data (the middle region), the cross-terms can produce large, negative signal due to couplings between $\textbf{x}$ and $\textbf{e}$ which affect 
$\widehat{\textbf{C}}_{r}$. This causes the difference between the green curve (which exhibits negligible loss at 
the data-only power spectrum value) and the black curve (which exhibits $\sim4$ orders of magnitude of loss). Right: The same power spectrum terms illustrated for the uniform weighted case.}
	\label{fig:sigloss_terms}
\end{figure*}

@@

The loss-corrected power spectrum limit for empirically estimated inverse covariance weighted PAPER-64 data is shown in Figure \ref{fig:ps2_data} (solid red), which we can compare to the original lossy result (dashed red). %There are a few important checks to examine. First, we see that the uniform-weighted power spectrum $2\sigma$ upper limit (dashed blue) is identical in both panels. This is an important check, as we expect no signal loss for this case. Any difference here would indicate an error in the pipeline or discrepancy in the range of data included. Additionally, it is clear that the power spectrum, prior to signal loss estimation, is inconsistent with the (now revised) theoretical noise level prediction. 
Post-signal loss estimation, the power spectrum limits are higher than both the theoretical noise level (green) and uniform-weighted power spectrum (which is shown three ways: black and gray points are positive and negative power spectrum values, respectively, with $2\sigma$ error bars from bootstrapping, the solid blue is the upper limit on the EoR signal using the full signal injection framework, and the shaded gray is the power spectrum values with thermal noise errors). We elaborate on this point in the next section, as well as investigate alternate 
weighting schemes to inverse covariance weighting, with the goal of finding one that balances the aggressiveness of down-weighting contaminants and minimizing the loss of the EoR signal.

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data.pdf}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data.pdf}
	\caption{A power spectrum of a subset of PAPER-64 data illustrating the use of empirical inverse covariance weighting. The solid red curve is the $2\sigma$ upper limit on the EoR signal estimated from our signal injection framework using empirical inverse covariance weighting. Shown for comparison is the lossy limit prior to signal loss estimation (dashed red). The theoretical $2\sigma$ thermal noise level prediction based on observational parameters is in green, whose calculation is detailed in Section \ref{sec:PSSense}. Additionally, the power spectrum results for the uniform weighted case is shown in three different ways: power spectrum values (black and gray points as positive and negative values, respectively, with $2\sigma$ error bars from bootstrapping), the $2\sigma$ upper limit on the EoR signal using our full signal injection framework (solid blue), and the measured power spectrum values with $2\sigma$ thermal noise errors (gray shaded regions). The vertical dashed black lines signify the horizon limit for this analysis using $30$\,m baselines. In this example, we see that the lossy power spectrum limit is $\sim 4$ orders of magnitude too low when using empirical inverse covariance weighting.}
\label{fig:ps2_data}
\end{figure*}

@@

We saw in Section \ref{sec:otherweight} how limiting the number of down-weighted eigenmodes (i.e. flattening out part of the 
eigenspectrum and effectively decoupling the lowest-valued eigenmodes, which are typically EoR-dominated, from the data) can help minimize signal loss. We experiment with this idea on PAPER-64 data, dialing the number of modes 
that are down-weighted from zero (which is equivalent to identity-weighting, or the uniform-weighted case) to $21$ (which is the full inverse 
covariance estimator). The power spectrum results for one $k$-value, both before and after signal loss 
estimation, are shown in the top panel in Figure \ref{fig:sigloss_modeloop}. We see that the amount of signal loss increases as weighting 
becomes more aggressive (dashed red). In other words, more EoR-dominated fluctuations are being overfit and 
subtracted as more modes are down-weighted. We also find that the power spectrum upper limit, post-signal loss estimation, 
increases with the number of down-weighted modes (solid red). The more modes we use in down-weighting, the stronger the coupling between the weighting and the data, and the greater the error we have in estimating the power spectrum. \citet{switzer_et_al2013} took a similar approach in determining the optimal number of modes to down-weight in GBT data, finding similar trends and noting that removing too few modes is limited by residual foregrounds and removing too many modes is limited by large error bars and signal loss.

Optimistically, we expect there to be a ``sweet spot" as we dial our regularization knob; a level of regularization where weighting 
is beneficial compared to uniform weighting (blue). In other words, we would like a weighting scheme that down-weights eigenmodes that predominantly describe foreground modes, but not EoR modes. We see in Figure \ref{fig:sigloss_modeloop} that this occurs roughly when 
only the $\sim2$-$3$ highest-valued eigenmodes are down-weighted and the rest are given equal weights (though for the case shown, weighting does not actually outperform uniform weighting). For a similar discussion on projecting out modes (zeroing out eigenmodes, rather than just ignoring their relative weightings as we do in this study), see \citet{switzer_et_al2013}.

We also saw in Section \ref{sec:otherweight} how adding the identity matrix to the empirical covariance can minimize signal loss. We experiment with this idea as well, shown in the bottom panel of Figure \ref{fig:sigloss_modeloop}. The dashed red and solid red lines represent power spectrum limits pre and post-signal loss estimation, respectively, as a function of the strength of $\textbf{I}$ that is added to $\widehat{\textbf{C}}$, quantified as a percentage of Tr($\widehat{\textbf{C}})\textbf{I}$ added to $\widehat{\textbf{C}}$. We parameterize this ``regularization strength" parameter as $\gamma$, namely $\widehat{\textbf{C}} \equiv \widehat{\textbf{C}} + \gamma$Tr$(\widehat{\textbf{C}})\textbf{I}$. From this plot we see that only a small percentage of Tr($\widehat{\textbf{C}})$ is needed to significantly reduce loss. We expect that as the strength of $\textbf{I}$ is increased (going to the left), both the red curves will approach the uniform-weighted case. We also notice that the post-signal loss limit hovers around the uniform-weighted limit for a large range of regularization strengths and while an overall trend from high-to-low signal loss is seen as the strength increases, there does not appear to be a clear ``minimum" that produces the least loss.

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_modeloop_2panel.pdf}
	\caption{Power spectra $2\sigma$ upper limits for $k=0.393$\,$h$ Mpc$^{-1}$ for fringe-rate filtered PAPER-64 data. Top: Values 
are shown before (dashed red) and after (solid red) signal loss estimation via our signal injection framework as a function of number of eigenmodes of $\widehat{\textbf{C}}$ that 
are down-weighted. This regularization knob is tuned from $0$ modes on the left (i.e. unweighted) to $21$ modes on the right (i.e. the full inverse 
covariance estimator). $\sim4$ orders of magnitude of signal loss results when using empirically estimated inverse covariance weighting. Bottom: Power spectrum upper limits before (dashed red) and after (solid red) signal loss estimation as a function of identity added to the empirical covariance. This regularization knob is tuned from $\gamma = 10^{-4}$ on the right (i.e. very little regularization) to $\gamma = 1$ on the left (see main text for the definition of $\gamma$). Also 
plotted in both panels for comparison are $2\sigma$ power spectrum upper limits for the uniform-weighted case (blue) and inverse variance 
weighted case (black); both are after signal loss estimation. Finally, a theoretical prediction for noise ($2\sigma$ error) is plotted 
as green. In the PAPER-64 analysis in this paper, we choose to use a regularization scheme of $\widehat{\textbf{C}}_{\rm eff} \equiv 0.09 \, $Tr($\widehat{\textbf{C}})\textbf{I} + \widehat{\textbf{C}}$ ($\gamma = 0.09$) as a simple example of regularization that minimizes loss, and note that the power spectrum limits using this type of regularization are roughly constant across a large range of values of $\gamma$.}
	\label{fig:sigloss_modeloop}
\end{figure*}

@@

The power spectrum result for our subset of PAPER-64 data (using only one baseline separation type, $10$ baselines, and $\widehat{\textbf{C}}_{\rm eff}$) is shown in Figure 
\ref{fig:ps1_data}. Again, the solid red curve represents our upper limit on the EoR signal using the full signal injection framework. The uniform weighted case is shown as the black and gray points, which correspond to positive and negative power spectrum values respectively (with 
$2\sigma$ errors bars from bootstrapping). It is also shown as an upper limit using the signal injection framework (solid blue), which is interestingly larger than the errors computed from bootstrapping, likely because the full injection framework takes into account additional sample variance whereas the bootstrapped errors do not. Finally, the gray shaded regions combine the measured uniform weighted power spectrum values with thermal noise errors. We show this power spectrum result as one example of how a simple regularization of an empirical covariance matrix can minimize signal loss, though we also note that this weighting does not produce more stringent limits than the uniform weighted case.
